A Model to Predict Survival in Patients with End-Stage Liver Disease

### 2021-08-07 analysis/article.Rmd
fix: observed vs expected mortality table

According to @wiesner2003 we divided our cohort into
`r nlevels(labelled_meld_data$MeldCategory)` MELD-score risk
categories, with `r enum(table(labelled_meld_data$MeldCategory))` patients,
respectively.
As depicted in Figure \@ref(fig:observed-vs-meld-mortality-plot) and Table
\@ref(tab:tbl-observed-vs-expected-mortality) our observed 90 day mortality was
`r paste(round(range(tbl_observed_vs_expected_mortality$SMR[-1]), 1), collapse = " to ")`
times higher than the predicted one. Just in the group with the lowest MELD
scores the mortality was much lower.

```{r tbl-observed-vs-expected-mortality, results = "asis", echo = FALSE, message = FALSE}
knitr::kable(
    cbind(
        row.names(tbl_observed_vs_expected_mortality),
        tbl_observed_vs_expected_mortality
    ),
    row.names = FALSE,
    col.names =
        c(
            "MELD category",
            "Observed deaths (n)",
            "Expected deaths (n)",
            "Standardized mortality ratio (SMR)",
            "Observed mortality (%)",
            "Expected mortality (%)"
        ),
    caption = paste0(
        "Observed vs MELD-expected 90 day mortality. ",
        "MELD mortality values are taken from @wiesner2003. ",
        "All patients censored before day 90 are ignored for ",
        " the calculation of the MELD-expected deaths. ",
        "SMR, Standardized mortality ratio = observed deaths/expected deaths."
    ),
    digits = 1
)
```

>>>>

According to @wiesner2003 we divided our cohort into
`r nlevels(labelled_meld_data$MeldCategory)` MELD-score risk
categories, with `r enum(table(labelled_meld_data$MeldCategory))` patients,
respectively.
As depicted in Figure \@ref(fig:observed-vs-meld-mortality-plot) and Table
\@ref(tab:tbl-observed-vs-expected-mortality) our observed 90 day mortality was
`r paste(round(range(tbl_observed_vs_expected_mortality$StandardizedMortalityRatio[-1]), 1), collapse = " to ")`
times higher than the predicted one. Just in the group with the lowest MELD
scores the mortality was much lower.

```{r tbl-observed-vs-expected-mortality, results = "asis", echo = FALSE, message = FALSE}
tbl_observed_vs_expected_mortality$ObservedMortality <-
    tbl_observed_vs_expected_mortality$ObservedMortality * 100
tbl_observed_vs_expected_mortality$ExpectedMortality <-
    tbl_observed_vs_expected_mortality$ExpectedMortality * 100
knitr::kable(
    cbind(
        row.names(tbl_observed_vs_expected_mortality),
        tbl_observed_vs_expected_mortality
    ),
    row.names = FALSE,
    col.names =
        c(
            "MELD category",
            "Observed deaths (n)",
            "Expected deaths (n)",
            "Standardized mortality ratio (SMR)",
            "Observed mortality (%)",
            "Expected mortality (%)"
        ),
    caption = paste0(
        "Observed vs MELD-expected 90 day mortality. ",
        "MELD mortality values are taken from @wiesner2003. ",
        "All patients censored before day 90 are ignored for ",
        " the calculation of the MELD-expected deaths. ",
        "SMR, Standardized mortality ratio = observed deaths/expected deaths."
    ),
    digits = 1
)
```
